---
title: "EDA on San Francisco Salaries"
author: "Aouidane Imed eddine"
output:
  pdf_document: default
  engine: xelatex
---
```{r}
library(readr)
library(tidyverse)
```


# About the dataset : 
- this data contains the names, job title, and compensation for San Francisco city employees on an annual basis from 2011 to 2014.
```{r,message=FALSE,warning=FALSE}
salaries <- read_csv("Salaries.csv")
dim(salaries)
```
```{r}
head(salaries)
```
- The dataset contains 13 columns and 148654 rows. The columns are:
  1. Id
  2. EmployeeName
  3. JobTitle
  4. BasePay
  5. OvertimePay
  6. OtherPay
  7. Benefits
  8. TotalPay
  9. TotalPayBenefits
  10. Year
  11. Notes
  12. Agency
  13. Status
  
# Removing unnecessary columns:
```{r}
salaries <- salaries %>% 
  select(-c(Notes, Agency,Id,EmployeeName,Year))
```


# Exploratory Data Analysis:

```{r}
summary(salaries)
```
- We can observe that the minimum salary is negative which is not possible. So we need to clean the data.
# Data Cleaning:
```{r}
salaries <- salaries %>% filter(BasePay > 0,
                                OvertimePay > 0,
                                OtherPay > 0,
                                Benefits > 0,
                                TotalPay > 0,
                                TotalPayBenefits >0)
summary(salaries)
```
- Now we have cleaned the data and removed the negative values. We can explore the data further.
## Full time jobs vs Part time jobs:

```{r}
# Full time jobs
summary(subset(salaries,Status == "FT")$TotalPay)
```
```{r}
# Part time jobs
summary(subset(salaries,Status == "PT")$TotalPay)
```
- The average salary for full-time jobs is 107,000 and for part-time jobs is 31,000. This shows that full-time jobs have higher salaries compared to part-time jobs.

## Visualizing the data:
```{r}
salaries_ft <- salaries[which(salaries$Status == "FT"),]
salaries_pt <- salaries[which(salaries$Status == "PT"),]
cbind(dim(salaries_ft),dim(salaries_pt))
```

### Full time jobs distribution:
```{r}
salaries_ft %>% ggplot(.,aes(x = TotalPay/1000))+
  geom_histogram(fill = "blue",bins = 100,,color = "white")+
  labs(title = "Full time jobs distribution",
       x = "Total Pay in thousands",
       y = "Frequency")
```
- The distribution of full-time jobs is right-skewed. Most of the employees earn between 0 to 200,000.

```{r}
salaries_ft %>% ggplot(.,aes(y = TotalPay/1000))+
  geom_boxplot(fill = "blue")+
  labs(title = "Full time jobs Boxplot",
       x = "Total Pay in thousands",
       y = "Frequency")
```
### Part time jobs distribution:
```{r}
salaries_pt %>% ggplot(.,aes(x = TotalPay/1000))+
  geom_histogram(fill = "red",bins = 100,color = "white")+
  labs(title = "Part time jobs distribution",
       x = "Total Pay in thousands",
       y = "Frequency")
```
- The distribution of part-time jobs is right-skewed. Most of the employees earn between 0 to 100,000.

```{r}
salaries_pt %>% ggplot(.,aes(y = TotalPay/1000))+
  geom_boxplot(fill = "red")+
  labs(title = "Part time jobs Boxplot",
       x = "Total Pay in thousands",
       y = "Frequency")
```
### Handeling outliers:
```{r}
salaries_ft <- salaries_ft %>% filter(TotalPay < mean(TotalPay) + 3*sd(TotalPay),
                                      TotalPay > mean(TotalPay) - 3*sd(TotalPay))

salaries_pt <- salaries_pt %>% filter(TotalPay < mean(TotalPay) + 3*sd(TotalPay),
                                      TotalPay > mean(TotalPay) - 3*sd(TotalPay))
```


### Comparing full-time and part-time jobs:

```{r}
salaries <- salaries %>% filter(!is.na(Status))
salaries %>% 
  ggplot(.,aes(x = TotalPay/1000,fill = Status))+
  geom_histogram(bins = 100,color = "white",alpha = 0.5)+
  labs(title = "Full time vs Part time jobs",
       x = "Total Pay in thousands",
       y = "Frequency")
```
### Density plot:
```{r}
salaries %>% 
  ggplot(.,aes(x = TotalPay/1000,fill = Status))+
  geom_density(alpha = 0.5)+
  labs(title = "Full time vs Part time jobs",
       x = "Total Pay in thousands",
       y = "Density")
```

## Conclusion:
- The average salary for full-time jobs is 107,000 and for part-time jobs is 31,000. This shows that full-time jobs have higher salaries compared to part-time jobs.


















